---
title: "Board Games"
date: "2019-03-12"
output:
  html_document:  
    keep_md: true
    fig_height: 6
    fig_width: 12
    fig_align: 'center'
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Data Exploration Workflow

Load in the usual packages, and ggthemes so I can use the fivethrityeight ggplot2 theme.
```{r message=FALSE}
library(tidyverse)
library(lubridate)
library(ggthemes)

board_games <- read_csv("https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2019/2019-03-12/board_games.csv")
```

Let's get a look at the data.

```{r}
glimpse(board_games)
```

First off, I want to see the cumulative board game total count over the years.

```{r}
cumulative <- 
  board_games %>% 
  count(year_published) %>% 
  mutate(cumulative = cumsum(n))
```

```{r}
cumulative %>% 
  ggplot(aes(year_published, cumulative)) +
  geom_point() +
  geom_line() +
  scale_x_continuous(breaks = seq(1950, 2020, 10)) +
  scale_y_continuous(labels = scales::comma) +
  labs(x = "Year",
       y = "Cumulative No. of Board Games") +
  theme_fivethirtyeight()
```
  
Looks like it started to pick up in the 70s and really got going exponential after 2000.

I want to do the same thing, but look at how the use of different game mechanics has changed over the years. A game mechanic is how a game is played like rolling dice, drawing, or storytelling. Let's take a look at the `mechanic` column.

```{r}
board_games %>% select(mechanic)
```

It looks like a board game can have multiple mechanic categories as we see here seperated by commas. What I need to do to analyze this type of data is force each row to have just one category while maintaining some contextual information. In this case I just want to keep a hold of the `year` column. Before we start, I'm just curious, what game had the most combinations of game mechanics?

```{r}
board_games %>% 
  select(year_published, name, mechanic) %>% 
  mutate(m_count = str_count(mechanic, ",")) %>%
  arrange(desc(m_count))
```

504. Hmm, I'm not familier with it. Looks like settlers of catan with a cash system after looking [here](https://boardgamegeek.com/image/2703890/504). Probably very complicated.

<br>

Now on to the data manipulation. I'll use `strsplit()` in a `mutate()` to serparate the categories and puts each of them into a nested tibble.

```{r}
mechanics_count <- 
  board_games %>% 
  select(year_published, mechanic) %>% 
  drop_na(mechanic) %>% 
  mutate(mechanic = strsplit(mechanic, ","))

mechanics_count
```

To make this nested data useful, we simply use `unnest()`

```{r}
mechanics_count <- 
  mechanics_count %>% 
  unnest()

mechanics_count
```

Now we have all the game mechanics used with their corresponding year. Notice we went from 9.5K rows to 23.9K which is expected.

```{r}
mechanics_count %>%
  select(mechanic) %>% 
  unique() %>%
  drop_na(mechanic) %>%
  nrow()
```

Before we get into plotting, I need to find the top 6 most occuring game mechanics. Having a total of 51 different game mechanics would make my plot hard to understand so let's keep it simple. 

```{r}
top_mechanics <- 
  mechanics_count %>% 
  count(mechanic, sort = T) %>% 
  top_n(n = 6, wt = n) %>% 
  pull(mechanic)

top_mechanics
```

Perfect, now lets plot them using a faceted density plot.

```{r}
mechanics_count %>% 
  filter(mechanic %in% top_mechanics) %>% 
  mutate(mechanic = factor(mechanic, levels = top_mechanics)) %>% 
  ggplot(aes(year_published, fill = mechanic)) +
  geom_density(col = "grey90", show.legend = F) +
  facet_wrap(~mechanic, ncol = 3) +
  scale_x_continuous(breaks = seq(1960, 2010, 10)) +
  scale_fill_tableau() +
  labs(title = "Most Occuring Board Game Mechanics over the years") +
  theme_fivethirtyeight() +
  theme(axis.text.y = element_blank(),
        axis.text.x = element_text(angle = 90))
```

All but one game mechanic seem to have rising in popularity rather quickly over the past two decades, while the Hex-and-Counter method has maintained a good share for for almost 40 years! 

@WargamesToGo from twitter asked, "Why do all of the curves drop off at the end. Isn't something rising in its place?" Interesting question! It does look like all six of the game mechanics in this graph decrease sharply before the hard cut off point in 2016. Here's some ideas of what may be happening.

- Perhaps there was an increase of video game popularity in the 2000s 
- @JoeWasserman suspects that the decreases are mostly an artifact or a flaw in the data collection. The final years "likely represent an incomplete record of games published."
- Or it very well could be other very new board game mechanics becomming more popular.

_*Note:* When @thomas_mock curated these data, he only included 12% of more than 90,000 games in the Board Game Geek database._

> To follow along with a fivethirtyeight article, I limited to only games with at least 50 ratings and for games between 1950 and 2016. This still leaves us with 10,532 games!

Maybe this removed more recent board games since they haven't had the same opportunity to be reviewed compared to older games.

Lets find out.

```{r}
library(bggAnalysis)
data("BoardGames")

board_games_full <- 
  BoardGames %>% 
  as_tibble() %>% 
  janitor::clean_names(.) %>% 
  select(year_published = details_yearpublished, 
         mechanic       = attributes_boardgamemechanic)

nrow(board_games_full)
```

Got the full data set! Now lets do the same thing we did before and see if anything changes.

```{r}
mechanics_long_full <- 
  board_games_full %>% 
  drop_na(mechanic) %>% 
  separate_rows(mechanic, sep = ",")

top_mechanics_full <- 
  mechanics_long_full %>% 
  count(mechanic, sort = T) %>% 
  top_n(n = 6, wt = n) %>% 
  pull(mechanic)

mechanics_long_full %>% 
  filter(mechanic %in% top_mechanics_full,
         year_published >= 1950,
         year_published <= 2016) %>% 
  mutate(mechanic = factor(mechanic, levels = top_mechanics_full)) %>% 
  ggplot(aes(year_published, fill = mechanic)) +
  geom_density(col = "grey90", show.legend = F) +
  facet_wrap(~mechanic, ncol = 3) +
  scale_x_continuous(breaks = seq(1960, 2010, 10)) +
  scale_fill_tableau() +
  labs(title = "Most Occuring Board Game Mechanics over the years") +
  theme_fivethirtyeight() +
  theme(axis.text.y = element_blank(),
        axis.text.x = element_text(angle = 90))

```



<br>

